---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# convmoment

<!-- badges: start -->

[![R-CMD-check](https://github.com/mahmudstat/convmoment/actions/workflows/pkgdown.yml/badge.svg)](https://github.com/mahmudstat/convmoment/actions/workflows/pkgdown.yml)
[![CRAN status](https://www.r-pkg.org/badges/version/convmoment)](https://CRAN.R-project.org/package=convmoment)
[![Documentation](https://img.shields.io/badge/docs-pkgdown-blue.svg)](https://www.thinkermahmud.com/convmoment/)
<!-- badges: end -->

**convmoment** implements the generalized binomial framework for transforming statistical moments of any order across arbitrary origins (Mahmud, 2025). It provides a unified, exact, and computationally efficient $O(r^2)$ method for:

* **Raw-to-Raw** transformations across arbitrary origins $a \to k$
* **Raw-to-Central** transformations (moments about origin $a$ to central moments about the mean)
* **Central-to-Raw** transformations (central moments back to raw moments about $0$)
* **Direct Moment Computation** from sample data up to degree $r$

The underlying mathematical identity uses the symbolic binomial operator:

$$ \mu'_r(k) \equiv (a + b)^r = \sum_{j=0}^r \binom{r}{j} \mu'_j(a) \, b^{r-j} $$

where $b = a - k$ is the origin-shift parameter and $a^j \mapsto \mu'_j(a)$.

---

## Installation

You can install the released version of **convmoment** from [CRAN](https://CRAN.R-project.org) (once accepted) with:

``` r
install.packages("convmoment")
```

Or install the development version from [GitHub](https://github.com/mahmudstat/convmoment) using `pak`:

``` r
# install.packages("pak")
pak::pak("mahmudstat/convmoment")
```

---

## Key Functions

| Function | Description |
| :--- | :--- |
| `conv_moment_all(x, a, k)` | Convert a vector of raw moments (orders $1 \dots K$) from origin $a$ to target origin $k$ |
| `raw2central(raw_moments, origin)` | Convert raw moments about `origin` to central moments (about the mean) |
| `central2raw(central_moments, mean_val)` | Convert central moments to raw moments about origin $0$ |
| `get_moments(x, a, r, na_rm, decimal)` | Compute raw or central moments directly from sample data $x$ up to degree $r$ |

---

## Examples

### 1. Raw-to-Raw Transformation Across Origins

Suppose the raw moments of a variable about origin $a = 2$ are $\mu'_1(2) = 1$, $\mu'_2(2) = 16$, and $\mu'_3(2) = -40$. What are the raw moments about origin $k = 0$?

```{r raw-to-raw}
library(convmoment)

# Raw moments about origin a = 2
x <- c(1, 16, -40)

# Convert to origin k = 0
conv_moment_all(x, a = 2, k = 0)
```

---

### 2. Raw-to-Central and Central-to-Raw Conversions

Convert raw moments about 0 to central moments (about the mean) and perform a round-trip check:

```{r raw-central}
set.seed(42)
data <- rnorm(200, mean = 10, sd = 3)

# Compute raw moments about 0 (orders 1..5)
raw_m <- sapply(1:5, function(r) mean(data^r))

# Convert raw moments to central moments
central_m <- raw2central(raw_m, origin = 0)
central_m

# Round-trip back to raw moments
raw_reconstructed <- central2raw(central_m, mean_val = mean(data))
all.equal(raw_m, raw_reconstructed)
```

---

### 3. Compute Moments Directly from Sample Data

Calculate raw moments around origin $a = 2$ up to 4th order:

```{r get-moments}
set.seed(123)
sample_data <- rnorm(100, mean = 5, sd = 2)

# Compute moments about origin a = 2 up to order r = 4
get_moments(sample_data, a = 2, r = 4, decimal = 4)
```

---

## Documentation & Vignette

For full theoretical details, high-order examples, distribution benchmarks (Binomial, Normal), and matrix formulations, visit the online documentation website:

👉 **[https://www.thinkermahmud.com/convmoment/](https://www.thinkermahmud.com/convmoment/)**

---

## Citation

If you use `convmoment` in your work, please cite:

```text
Mahmud, A. A. (2025). A Generalized Binomial Framework for Moment Transformations 
Across Arbitrary Origins. R package version 0.2.3.
```

---

## License

This package is licensed under the [MIT License](LICENSE.md).
