%\VignetteIndexEntry{Generalized Binomial Moment Transformations}
%\VignetteEngine{knitr::rmarkdown}
%\VignetteEncoding{UTF-8}
---
title: "Generalized Binomial Moment Transformations with convmoment"
author: "Abdullah Al Mahmud"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generalized Binomial Moment Transformations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(convmoment)
```

## Introduction

The `convmoment` package implements the **generalized binomial framework** for transforming statistical moments across arbitrary origins, as described in Mahmud (2025). This framework unifies raw-to-raw, raw-to-central, and central-to-raw moment conversions under a single mathematical principle:

$$ \mu'_r(k) = \sum_{j=0}^r \binom{r}{j} \mu'_j(a) \, b^{r-j} $$

where $b = a - k$ is the origin shift parameter.

The key insight is the **symbolic binomial operator**:

$$ \mu'_r(k) \equiv (a + b)^r \quad \text{with} \quad a^j \mapsto \mu'_j(a) $$

This allows any moment transformation to be constructed instantly using Pascal's triangle coefficients.

## Installation

```r
# From GitHub (once published)
# remotes::install_github("yourusername/convmoment")

# From local source
# remotes::install_local("path/to/convmoment")
library(convmoment)
```

## Core Functions

| Function | Purpose |
|----------|---------|
| `conv_moment_all(x, a, k)` | Convert all moments up to order K from origin `a` to origin `k` |
| `raw2central(raw_moments, origin)` | Convert raw moments about `origin` to central moments |
| `central2raw(central_moments, mean_val)` | Convert central moments to raw moments about 0 |
| `get_moments(x, a, r)` | Compute raw/central moments directly from data |

## Basic Usage: Raw-to-Raw Transformation

```{r raw-to-raw}
# Example from the paper: moments about a=2 to moments about k=5
# Given: mu'_1(2) = -1, mu'_2(2) = 7, mu'_3(2) = 39
moments_a2 <- c(-1, 7, 39)

# Convert to origin k=5 (b = a - k = 2 - 5 = -3)
moments_k5 <- conv_moment_all(moments_a2, a = 2, k = 5)
moments_k5

# Expected: mu'_1(5) = -4, mu'_2(5) = 22, mu'_3(5) = -78
```

## Raw-to-Central Transformation

```{r raw-to-central}
# Generate sample data
set.seed(42)
x <- rnorm(200, mean = 10, sd = 3)

# Compute raw moments about origin 0 (orders 1-6)
raw_moments <- sapply(1:6, function(r) mean(x^r))

# Convert to central moments
central_moments <- raw2central(raw_moments, origin = 0)

# Verify against direct computation
direct_central <- sapply(1:6, function(r) mean((x - mean(x))^r))

# Compare
results <- data.frame(
  Order = 1:6,
  Direct = direct_central,
  Converted = central_moments,
  Difference = abs(direct_central - central_moments)
)
knitr::kable(results, digits = 10,
  caption = "Raw-to-Central: Converted vs Direct Computation")
```

## Central-to-Raw Transformation

```{r central-to-raw}
# Using the central moments from above
mean_x <- mean(x)
raw_back <- central2raw(central_moments, mean_val = mean_x)

# Verify against original raw moments
results2 <- data.frame(
  Order = 1:6,
  Original = raw_moments,
  Reconstructed = raw_back,
  Difference = abs(raw_moments - raw_back)
)
knitr::kable(results2, digits = 10,
  caption = "Central-to-Raw: Reconstructed vs Original")
```

## Round-Trip Validation

```{r round-trip}
# Full cycle: raw -> central -> raw
raw_reconstructed <- central2raw(
  raw2central(raw_moments, origin = 0),
  mean_val = mean_x
)

max_diff <- max(abs(raw_moments - raw_reconstructed))
max_diff
```

The round-trip error is on the order of machine precision (~1e-15), confirming numerical exactness.

## High-Order Moments

The framework scales naturally to arbitrary high orders without manual formula derivation:

```{r high-order}
# 10th order moment transformation
set.seed(123)
x <- rnorm(1000, mean = 5, sd = 2)
raw_10 <- sapply(1:10, function(r) mean(x^r))
central_10_direct <- sapply(1:10, function(r) mean((x - mean(x))^r))
central_10_converted <- raw2central(raw_10, origin = 0)

high_order_results <- data.frame(
  Order = 1:10,
  Direct = central_10_direct,
  Converted = central_10_converted,
  Diff = abs(central_10_direct - central_10_converted)
)
knitr::kable(high_order_results, digits = 8)
```

## Computational Performance

Transforming pre-computed moments is **sample-size independent** ($O(r^2)$) versus recomputing from raw data ($O(n \cdot r)$):

```{r benchmark}
set.seed(42)
n <- 100000
x_large <- rnorm(n, mean = 10, sd = 2)
origin_a <- 2
origin_k <- 5

# Pre-computed moments about origin_a
raw_a <- sapply(1:5, function(r) mean((x_large - origin_a)^r))

N_calls <- 1000

# Benchmark 1: Recompute from raw data (O(n * r))
t_raw <- system.time({
  for (i in 1:N_calls) {
    dummy <- sapply(1:5, function(r) mean((x_large - origin_k)^r))
  }
})["elapsed"]

# Benchmark 2: Binomial transformation (O(r^2))
t_binom <- system.time({
  for (i in 1:N_calls) {
    dummy <- conv_moment_all(raw_a, a = origin_a, k = origin_k)
  }
})["elapsed"]

speedup <- round(t_raw / t_binom, 1)

bench <- data.frame(
  Method = c("Raw Data Recomputation (O(n*r))", "Binomial Transformation (O(r^2))"),
  Time_sec = c(t_raw, t_binom),
  Speedup = c("1.0x (baseline)", paste0(speedup, "x faster"))
)
knitr::kable(bench, digits = 3,
  caption = "Performance: 1,000 transformations on n = 100,000")
```

## Theoretical Distributions

The framework works directly with theoretical moments from named distributions:

### Binomial Distribution

```{r binomial}
# Binomial(n=10, p=0.3): theoretical raw moments about 0
n <- 10; p <- 0.3
# Using known formulas or MGF derivatives
mu1_0 <- n * p
mu2_0 <- n * p * (1 - p) + n^2 * p^2
mu3_0 <- n * p * (1 - p) * (1 - 2 * p) + 3 * n^2 * p^2 * (1 - p) + n^3 * p^3
mu4_0 <- n * p * (1 - p) * (1 - 6 * p * (1 - p)) +
         6 * n^2 * p^2 * (1 - p) * (1 - 2 * p) +
         7 * n^3 * p^3 * (1 - p) + n^4 * p^4

raw_binom <- c(mu1_0, mu2_0, mu3_0, mu4_0)
central_binom <- raw2central(raw_binom, origin = 0)

# Known theoretical central moments:
# mu_1 = 0
# mu_2 = n*p*(1-p) = 2.1
# mu_3 = n*p*(1-p)*(1-2p) = 0.84
# mu_4 = 3*n^2*p^2*(1-p)^2 + n*p*(1-p)*(1-6*p*(1-p)) = 3.294 + 0.441 = 3.735

theoretical <- c(0, n*p*(1-p), n*p*(1-p)*(1-2*p),
  3*n^2*p^2*(1-p)^2 + n*p*(1-p)*(1-6*p*(1-p)))

results_binom <- data.frame(
  Order = 1:4,
  Converted = central_binom,
  Theoretical = theoretical,
  Diff = abs(central_binom - theoretical)
)
knitr::kable(results_binom, digits = 8)
```

### Normal Distribution

```{r normal}
# Normal(mu=2, sigma^2=4): theoretical raw moments about 0
mu <- 2; sigma2 <- 4
raw_normal <- c(
  mu,
  mu^2 + sigma2,
  mu^3 + 3*mu*sigma2,
  mu^4 + 6*mu^2*sigma2 + 3*sigma2^2
)
central_normal <- raw2central(raw_normal, origin = 0)

# Theoretical central moments of Normal:
# mu_1 = 0, mu_2 = sigma^2 = 4, mu_3 = 0, mu_4 = 3*sigma^4 = 48
theoretical_normal <- c(0, sigma2, 0, 3*sigma2^2)

results_normal <- data.frame(
  Order = 1:4,
  Converted = central_normal,
  Theoretical = theoretical_normal,
  Diff = abs(central_normal - theoretical_normal)
)
knitr::kable(results_normal, digits = 8)
```

## Matrix Formulation

For computational implementations, the transformation can be expressed as a matrix-vector product:

```{r matrix-form}
# Transformation matrix C_b for orders 0..K
K <- 5
b <- 2 - 5  # = -3
C_b <- outer(0:K, 0:K, function(r, j) ifelse(j <= r, choose(r, j) * b^(r-j), 0))
C_b

# Verify: mu'(k) = C_b %*% mu'(a)
mu_a <- c(1, -1, 7, 39, NA, NA)  # mu_0=1, mu_1=-1, mu_2=7, mu_3=39
mu_a[5:6] <- c(
  conv_moment(c(-1,7,39,NA,NA), 2, 5, 4),  # computed
  conv_moment(c(-1,7,39,NA,NA), 2, 5, 5)
)
C_b %*% mu_a
```

## References

- Mahmud, A. A. (2025). *A Generalized Binomial Framework for Moment Transformations Across Arbitrary Origins*. [Journal submission].
- Pébay, P. (2008). *Formulas for robust, one-pass parallel computation of covariances and arbitrary-order statistical moments*. Sandia National Laboratories Technical Report SAND2008-6212.